Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint Estimation

### 2015-04-08 2349.tex
CVPR cam-ready. reduced the size of images, major updates on the extended cam-ready

% custom commands
\newcommand{\scream}[1]{{\color{red} \bf *** #1 ***}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{1}
\begin{document}

@@

\maketitle
%\thispagestyle{empty}

@@

\paragraph{Acknowledgement}
This work has been supported by the Max Planck Center for Visual Computing \& Communication.

>>>>

% custom commands
\newcommand{\scream}[1]{{\color{red} \bf *** #1 ***}}
% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
% \setcounter{page}{1}
\begin{document}

@@

\maketitle
\thispagestyle{empty}

@@

\paragraph{Acknowledgement}
We acknowledge the support of NSF CAREER grant (N1054127), Ford-Stanford
Innovation Alliance Award, DARPA, Korea Foundation for Advanced Studies and
the Max Planck Center for Visual Computing \& Communication.

### 2015-04-08 2349_ext.tex
CVPR cam-ready. reduced the size of images, major updates on the extended cam-ready

\noindent
% In this paper, we propose a novel method for 2D-3D alignment of exemplar CAD
% models to real-world images that circumvents the need for calibration and
% greatly enhances the scalability of WHO. Our model generates a discriminative
% exemplar template on-the-fly and align model accurately using MCMC.

A large body of recent work on object detection has focused on providing more
information than a bounding box of an object. Some used 3D CAD
model databases to provide viewpoint estimation. These approaches work
by aligning exact 3D models to images using templates generated from renderings
of the 3D models at a set of discrete viewpoints.

@@

\begin{figure}[t]
  \centering
  % \includegraphics[width=0.9\linewidth]{schematics} 
  \includegraphics[width=0.9\linewidth]{front} % \\[-5pt]
  \caption{Using a database of 3D CAD models, we generate NZ-WHO
    templates which can be used to either detect objects directly or
    enrich the output of an existing detector with high-quality,
    continuous pose and 3D CAD model exemplar.}
    % initialize pose of an object in the bounding box from other
    % detectors. Once initialization is given, we use NZ-WHO to propose
    % and validate plausible pose on-the-fly and
    % further tune the translation, 3D viewpoint and focal length continuously.}
  % \vspace{-1em}
  \label{fig:front}
\end{figure}

\vspace{-0.1in}
\paragraph{Overview}

In this paper, we propose a novel method for 2D-3D alignment of exemplar CAD
models to real-world images that circumvents the need for calibration and
greatly enhances the scalability of WHO. As a result, we can render novel views
and train corresponding exemplar models {\em on-the-fly}, without the need for
offline processing. We call these Non-Zero Whitened Histograms of Orientations
(NZ-WHO) templates. First, we generate an ensemble of NZ-WHO templates to find
coarse viewpoints and fine-tune them using MCMC. The schematic of our pipeline
is on Fig.~\ref{fig:front}. As a by-product, we can formulate the alignment
problem as a parameter search in a continuous pose space, consisting of yaw,
pitch and roll, which we implement using MCMC sampling.

First, we present a novel method for training exemplar models from
rendered 3D CAD data on-the-fly, enabled by a novel variant of
WHO, termed Non-Zero Whitened Histograms of Orientations (NZ-WHO), and making
efficient use of the specific characteristics of rendered images. To our
knowledge, our method constitutes the first attempt to simultaneously render
and train exemplar detectors on-the-fly.

Second, we demonstrate that our method can enrich the output of an existing
object class detector, such as the DPM or the
R-CNN with additional 3D information. By applying our
method to candidate detections provided by the respective detector, we augment the original detections with an estimate of 3D continuous pose and a 3D
CAD model exemplar.

Finally, our in-depth experimental study demonstrates
the effectiveness of the approach on a standard benchmark for
object detection and viewpoint estimation~\cite{Xiang14}.

@@

\vspace{-0.1in}
\paragraph{Non-Zero Whitened Histograms of Orientations}

NZ-WHO removes these artifacts so that the background has no effect on the
template response. Let a vectorized HOG feature of a rendering image $x = [
x_1^T ,  \cdots , x_n^T ]^T \in \mathcal{R}^{nd}$ where $x_i$ is the $i$th HOG cell feature, $n$ is the
number of HOG cells and $d$ is the dimension of the HOG feature. We create a new
vector $\bar{x}$ which contains only the non-zero HOG cell features of $x$.
Synthesizing the LDA template requires solving the system of linear equations
$\bar{w}=\bar{\Sigma}^{-1}(\bar{x} - \bar{\mu})$. In \cite{Hariharan12}, the
authors solve the system via the Cholesky decomposition, which requires
$O(n^3)$ time. Instead, we introduced the Conjugate Gradient method to generate
LDA templates. The method runs in $O(n^2\kappa)$ time, where $\kappa$ is the condition number of the matrix and achieved more than 100x speed-up.

@@

\vspace{-0.1in}
\paragraph{Fine-Tuning using MCMC}
Due to the speed of our template synthesis procedure, we are able to perform
joint optimization of scale, translation, continuous rotation, and focal length
using the Metropolis-Hastings algorithm.

@@

\vspace{-0.1in}
\paragraph{Experiments}

First, we verify that our NZ-WHO method delivers performance that is
at least on par with the original WHO formulation~\cite{Hariharan12} in terms of
accuracy, while at the same time resulting in large computational
savings.

@@

>>>>

\begin{figure}[t]
  \centering
  % \includegraphics[width=0.9\linewidth]{schematics} 
  \includegraphics[width=0.9\linewidth]{front} % \\[-5pt]
  \caption{Using a database of 3D CAD models, we generate NZ-WHO
    templates which can be used to either detect objects directly or
    enrich the output of an existing detector with high-quality,
    continuous pose and 3D CAD model exemplar.}
    % initialize pose of an object in the bounding box from other
    % detectors. Once initialization is given, we use NZ-WHO to propose
    % and validate plausible pose on-the-fly and
    % further tune the translation, 3D viewpoint and focal length continuously.}
  % \vspace{-1em}
  \label{fig:front}
\end{figure}
\vspace{-0.15in}

% In this paper, we propose a novel method for 2D-3D alignment of exemplar CAD
% models to real-world images that circumvents the need for calibration and
% greatly enhances the scalability of WHO. Our model generates a discriminative
% exemplar template on-the-fly and align model accurately using MCMC.

\noindent
A large body of recent work on object detection has focused on providing more
information than a bounding box of an object. Some used 3D CAD
model databases to provide viewpoint estimation. These approaches work
by aligning exact 3D models to images using templates generated from renderings
of the 3D models at a set of discrete viewpoints~\cite{Aubry14}.

@@

\vspace{-0.15in}
\paragraph{Overview.} In this paper, we propose a novel method for 2D-3D
alignment of exemplar CAD models to real-world images that circumvents the need
for calibration and greatly enhances the scalability of WHO. As a result, we
can render novel views and train corresponding exemplar models {\em
	on-the-fly}, without the need for offline processing. We call these Non-Zero
Whitened Histograms of Orientations (NZ-WHO) templates. To our knowledge, our
method constitutes the first attempt to simultaneously render and train
exemplar detectors on-the-fly.

First, we adapt the whitening to the specific case of rendered images. We show
how to speed up the whitening by two orders of magnitude for high-resolution
templates using Conjugate Gradient method. Also, we improve the evaluation of
our 3D exemplar template detectors at test time by performing convolutions in
the frequency domain.

These components make template generation and evaluation computationally
inexpensive. Thus, we are able to efficiently explore the continuous parameter
space to find the best object pose, scale, 3D CAD model type and camera focal
length. The schematic of our pipeline is on Fig.~\ref{fig:front}.

Finally, our experimental study demonstrates the effectiveness of the
approach on several standard benchmarks for object detection and viewpoint
estimation. We also demonstrate that our method can enrich the output of an
existing object class detector, such as the DPM or the R-CNN with additional 3D
information.

@@

\vspace{-0.1in}
\paragraph{Non-Zero Whitened Histograms of Orientations.} Many works that use
rendering did not take the effect of background into account~\cite{Aubry14}
when synthesizing and whitening to generate an LDA template. Our NZ-WHO removes
background region while whitening thus it prevents irrelevant regions from
corrupting the foreground.

NZ-WHO method greatly speeds up the LDA generation. We introduced the Conjugate
Gradient method which runs in $O(n^2\kappa)$ time for LDA generation. Finally,
we experimentally show that our uncalibrated NZ-WHO template performs on par
with a calibrated WHO template but 100x faster to generate.

@@

\vspace{-0.1in}
\paragraph{Fine-Tuning using MCMC.} Due to the speed of our template synthesis
procedure, we are able to perform joint optimization of scale, translation,
continuous rotation, and focal length using the Metropolis-Hastings algorithm.

@@

\vspace{-0.1in}
\paragraph{Experiments} First, we verify that our NZ-WHO method delivers
performance that is at least on par with the original WHO
formulation~\cite{Hariharan12} in terms of accuracy, while at the same time
resulting in large computational savings.

@@

\vspace{-0.1in}
\paragraph{Acknowledgement.} We acknowledge the support of NSF CAREER grant
(N1054127), Ford-Stanford Innovation Alliance Award, DARPA, Korea Foundation
for Advanced Studies and the Max Planck Center for Visual Computing \&
Communication.
